---
title: "deep_learning_eda"
author: "Sreeram Peela"
date: "2024-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
The file contains exploratory data analysis (EDA) of "Churn modelling" dataset.
The dependent feature is "Exited", while all the other features are considered
features. Prior to usig the dataset for classification problems, I would like to
perform a biref statistical analysis.

The goal is to identify features that are significantly different in the groups:
exited (encoded as 1) and not-exited (encoded as 0). For the purpose of EDA, the
original datafile is used, and not the cleaned one (with scaled data).

```{r, echo=T}
## load required packages
library(ggplot2)
library(dplyr)
library(reshape2)
```

Loading the dataset
```{r}
df <- read.csv("Churn_Modelling.csv", stringsAsFactors = T)
```

Performing some feature recoding:
Some of the binary/categorical features are encoded as 0 and 1.
```{r, echo=T}
#removing customer id and surname features as they are not relevant
df <- select(df, -c(1,2))
summary(df)
```
```{r, echo=T}
#recoding as factors
df$NumOfProducts <- as.factor(df$NumOfProducts)
df$HasCrCard <- as.factor(df$HasCrCard)
df$IsActiveMember <- as.factor(df$IsActiveMember)
summary(df)
```
```{r, echo=T}
cat_results <- list(pvalues = c(), odds_ratio = c(), feature_names = c())
#using fisher test or chisquare test for categorical associations
cat_features <- c(2,3,7,8,9)
for (i in cat_features) {
  unique_cats <- unique(df[ , i])
  cross_tab <- table(df[ , i], df[ , "Exited"])
  colName <- colnames(df)[i]
  cat_results$feature_names <- c(cat_results$feature_names, colName )
  #cat_results$cont_tables <- c(cat_results$cont_tables, matrix(cross_tab,
                              #nrow=nrow(cross_tab), ncol = ncol(cross_tab)))
  if (length(unique_cats) == 2) {
    # run Fisher test
    fisher_test <- fisher.test(cross_tab, conf.int = TRUE, conf.level = 0.95)
    cat_results$pvalues <- c(cat_results$pvalues, fisher_test$p.value)
    cat_results$odds_ratio <- c(cat_results$odds_ratio, fisher_test$estimate)
    #cat_results$odd_ratio_CI <- c(cat_results$odd_ratio_CI, fisher_test$conf.int)
    
  } else {
    # run chisquare test
    # no odds ratio for chi-square test
    chi_test <- chisq.test(cross_tab)
    cat_results$pvalues <- c(cat_results$pvalues, chi_test$p.value)
    cat_results$odds_ratio <- c(cat_results$odds_ratio, NA)
    #cat_results$odd_ratio_CI <- c(cat_results$odd_ratio_CI, c(NA, NA))
    
  }
}
print(cat_results)
```
```{r, echo=T}
cat_results_df <- data.frame("feature" = cat_results$feature_names, 
                             "pval" = cat_results$pvalues,
                             "odds" = round(cat_results$odds_ratio, 3))
sign_features <- cat_results_df[cat_results_df$pval<0.05, "feature"]
print(sign_features)

```
From the above observation, has card doesn't seem to be associated with exited status.

Performing EDA using continuous variables

```{r, echo=T}
cont_features <- c(1,4,5,6,10)
shapiro_results <- list("pvalue" = c(), "feature" = c())
#checking normality assumptions
# using Anderson-Darling normality test as n_samples > 5000
library(nortest)
for (i in cont_features) {
  pval <- ad.test(df[ , i])$p.value
  colname <- colnames(df)[i]
  shapiro_results$pvalue <- c(shapiro_results$pvalue, pval)
  shapiro_results$feature <- c(shapiro_results$feature, colname)
}

print(shapiro_results)
  
```
As seen from above, none of the variables are normally distributed. Using correlation to detect collinear variables

```{r, echo=T}
df_cont <- select(df, cont_features)
cor_matrix <- cor(df_cont)
print(cor_matrix)
pairs(df_cont)
```
```{r, echo=T}
df_cont$Exited <- df$Exited
levels(df_cont$Exited)[levels(df_cont$Exited) == "1"] <- "YES"
levels(df_cont$Exited)[levels(df_cont$Exited) == "0"] <- "NO"

```


Using Mann-Whitney test for two groups comparison
As wilcox.test is the function used

```{r, echo=T}
pvals = c()
var_names = c()
for (i in 1:4) {
  varname = colnames(df_cont)[i]
  ind_var <- df_cont[ , i]
  stat_test <- wilcox.test(ind_var~df_cont$Exited)
  var_names <- c(var_names, varname)
  pvals <- c(pvals, stat_test$p.value)
  
}

cont_res <- list("pvals"=pvals, "feature" = var_names)

```


Generating box plots
```{r, echo=T}
# drawing boxplots
cont_vars <- colnames(df_cont)[1:4]
library(ggpubr)
library(gridGraphics)
# we have four continuous features
# creating a 2x2 plot
#par(mfrow=c(2,2))
plot1 <- boxplot(df$EstimatedSalary ~ df$Exited, xlab = "Exited", 
                 ylab = "Estimated salary")
plot2 <- boxplot(df$CreditScore ~ df$Exited, xlab = "Exited", 
                 ylab = "Credit score")
plot3 <- boxplot(df$Balance ~ df$Exited, xlab = "Exited", 
                 ylab = "Balance")
plot4 <- boxplot(df$Tenure ~ df$Exited, xlab = "Exited", 
                 ylab = "Tenure")
plot5 <- boxplot(df$Age ~ df$Exited, xlab = "Exited", 
                 ylab = "Age")
```
```{r, echo=T}
pvals = c()
var_names = c()
for (i in 1:4) {
  varname = colnames(df_cont)[i]
  ind_var <- df_cont[ , i]
  stat_test <- wilcox.test(ind_var~df_cont$Exited)
  var_names <- c(var_names, varname)
  pvals <- c(pvals, stat_test$p.value)
  
}

cont_res <- list("pvals"=pvals, "feature" = var_names)



```

```{r, echo=T}
cont_res_df <- data.frame("feature" = cont_res$feature, "pval" = cont_res$pvals)
cont_res_df[cont_res_df$pval < 0.05, "feature"]

```

Interpretation: As seen from the above results, only these three variables were 
significantly different among those exited and non-exited.
